\documentclass[11pt,a4paper]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{microtype}
\usepackage{array}
\usepackage{cite}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    citecolor=black,
    urlcolor=blue
}

\title{\textbf{Dragonfly Algorithm}\\
       \large Synopsis: CITS4404 Part 1 Literature Review}
\author{James Wigfield \\ student ID: 23334375}
\date{April 2026}

\begin{document}

\maketitle

\noindent\textbf{Source paper:} S.\ Mirjalili, ``Dragonfly algorithm: a new meta-heuristic optimization technique for solving single-objective, discrete, and multi-objective problems,'' \textit{Neural Computing and Applications}, vol.~27, pp.~1053--1073, 2016.~\cite{mirjalili2016dragonfly}

\bigskip

\section*{1.\ What problem is the algorithm attempting to solve?}

The Dragonfly Algorithm (DA) is a swarm intelligence (SI) optimiser proposed to address continuous, binary, and multi-objective optimisation problems.
Although previous algorithms such as the Particle Swarm Optimisation (PSO)~\cite{kennedy1995pso}, Ant Colony Optimisation (ACO), and Artificial Bee Colony (ABC) 
have been widely studied, they model only a subset of the behavioural repertoire available in nature.
Many swarming phenomena, dragonfly swarming in particular, had not been studied computationally.
The No Free Lunch (NFL) theorem~\cite{wolpert1997nfl} provides further motivation: no single algorithm is optimal across all problems, so a genuinely distinct algorithm can outperform existing ones on problems where they are weak.

\bigskip



\section*{2.\ Why have previous attempts failed?}

Prior SI algorithms capture different aspects of swarm behaviour but none models all five survival-critical behaviours simultaneously.
PSO, for example, uses only attraction towards a personal best and a global best (analogous to cohesion and food attraction), and several improved PSO variants have attempted to add separation, alignment, or cohesion operators independently~\cite{cui2009boid,kadrovach2002particle,cui2009alignment} but never integrated all five factors into a unified framework.
More fundamentally, existing algorithms do not distinguish between two qualitatively different modes of swarming: \textit{static} swarms that hunt over small areas (exploration) and \textit{dynamic} migratory swarms that travel in one direction (exploitation).
PSO blurs this distinction through a single velocity update rule, making the exploration/exploitation trade-off difficult to control directly.
The absence of an explicit \textit{enemy} (worst solution) in standard formulations also means search agents are only attracted toward promising regions but never actively repelled from non-promising ones, potentially reducing diversity in the early stages of search.

\bigskip

\section*{3.\ What is the new idea?}

DA models artificial dragonflies using \textbf{five behavioural operators} derived from Reynolds' flocking rules~\cite{reynolds1987flocks} augmented with food/enemy interactions:

\begin{enumerate}
    \item \textbf{Separation} ($S_i = -\sum_j (X - X_j)$): avoids collision with neighbours.
    \item \textbf{Alignment} ($A_i = \frac{1}{N}\sum_j V_j$): matches neighbours' velocities.
    \item \textbf{Cohesion} ($C_i = \frac{1}{N}\sum_j X_j - X$): moves toward the neighbourhood centre of mass.
    \item \textbf{Food attraction} ($F_i = X^+ - X$): steers toward the current best solution.
    \item \textbf{Enemy distraction} ($E_i = X^- + X$): steers away from the current worst solution.
\end{enumerate}

The \textbf{step vector} (analogous to PSO velocity) combines all five:
\[
    \Delta X_{t+1} = (s\,S_i + a\,A_i + c\,C_i + f\,F_i + e\,E_i) + w\,\Delta X_t
\]
where $w$ decays from 0.9 to 0.2 over iterations to shift from exploration to exploitation.
When a dragonfly has no neighbours it performs a \textbf{Lévy flight} ($X_{t+1} = X_t + \text{Levy}(d)\cdot X_t$), improving stochastic coverage of the search space.

Static swarming (high alignment weight $a$, low cohesion weight $c$) models exploration; dynamic swarming (low $a$, high $c$) models exploitation.
The neighbourhood radius grows monotonically so the population transitions from many small sub-swarms to one large convergent swarm over the course of optimisation.

Two variants are also presented.
The \textbf{Binary DA (BDA)} applies a v-shaped transfer function to convert continuous step values into flip probabilities, since v-shaped functions avoid the forcing of particles to extreme binary values that plagues s-shaped alternatives~\cite{saremi2014transfer}.
The \textbf{Multi-Objective DA (MODA)} adds a Pareto archive and a hypersphere grid that selects food sources from the least-populated grid cells (improving coverage) and enemies from the most-populated cells (discouraging revisiting crowded regions), following the MOPSO archive strategy~\cite{coello2004mopso}.

\bigskip

\section*{4.\ How is the new approach demonstrated?}

The paper validates DA, BDA, and MODA through three experiment sets.

\textbf{Single-objective DA:} 19 benchmark functions across three categories — 7 unimodal (TF1--TF7, testing exploitation/convergence), 6 multi-modal (TF8--TF13, testing exploration and local-optima avoidance), and 6 composite (TF14--TF19, shifted/rotated combinations mimicking real landscapes) — are evaluated using 30 search agents over 500 iterations, repeated 30 times.
Comparison algorithms are PSO (original parameters) and GA (original parameters).
Statistical significance is assessed via the Wilcoxon rank-sum test.

\textbf{Binary BDA:} The same 13 functions (TF1--TF13) encoded in 75 binary variables (5 dimensions × 15 bits each) are compared against Binary PSO (BPSO) and Binary GSA (BGSA).

\textbf{Multi-objective MODA:} Five ZDT benchmark functions (convex, non-convex, discontinuous, linear, and tri-objective fronts) and a real-world 20-variable \textit{submarine propeller design} problem are solved.
Coverage and convergence are measured via Inverse Generalised Distance (IGD) over 10 runs and compared against MOPSO and NSGA-II.
The real-world case study uses 200 agents over 300 iterations with a death-penalty constraint-handling scheme.

Sufficient detail is provided (equations, pseudocode, parameter values, benchmark function definitions in appendices) to support replication, and source code is publicly available.

\bigskip

\section*{5.\ Results and validation}

\textbf{DA vs PSO vs GA (single-objective):} DA achieves the best or joint-best result on 13 of 19 functions, and Wilcoxon p-values confirm statistical significance over GA on virtually all functions ($p < 0.0002$).
Against PSO, differences are often not statistically significant on multi-modal functions (e.g.\ TF9, TF11, TF12), suggesting comparable exploration ability.
A notable exception is TF8 (Schwefel), where PSO substantially outperforms DA; this is the only case where a single-function weakness appears systematic rather than marginal.
On composite functions (TF14--TF19) the advantage over PSO is less consistent, which the author attributes honestly to the difficulty of these landscapes.

\textbf{BDA:} Outperforms BPSO and BGSA on 11 of 13 binary functions, with most differences strongly significant ($p < 0.002$).

\textbf{MODA:} Consistently better than NSGA-II on all five ZDT variants by a large margin; competitive with MOPSO, outperforming it on ZDT3 and the three-objective variant.
For the propeller design problem, MODA produced 61 well-distributed Pareto optimal solutions under strong constraints, demonstrating practical utility despite the absence of a known true Pareto front for verification.

Four behavioural diagnostics (search history, trajectory, average fitness, convergence curve) on selected functions confirm monotonic fitness improvement and eventual convergence, providing qualitative support for the algorithm's theoretical properties.

\bigskip

\section*{6.\ Assessment of conclusions}

The authors' primary claims --- that DA is a competitive single-objective optimiser and that BDA and MODA outperform their respective comparators --- are largely substantiated by the evidence.
The use of the Wilcoxon test and 30-run averaging follows sound methodology, and the breadth of benchmarks is commendable.

However, several caveats deserve attention.
First, the comparison set is narrow: PSO and GA are the only single-objective baselines.
Contemporary algorithms such as Differential Evolution or Grey Wolf Optimiser (also authored by Mirjalili) are absent, which limits the strength of the superiority claim.
Second, results on composite functions are mixed and the statistical advantage over PSO is modest, suggesting that DA's exploration/exploitation balance may not consistently generalise to the highly deceptive landscapes most characteristic of real-world problems.
Third, the propeller design comparison lacks an independently known true Pareto front, so MODA's performance cannot be benchmarked against an absolute standard.

For the purposes of this assignment, DA is an attractive candidate: its five operators and explicit food/enemy mechanism translate naturally to a continuous search space such as the trading-bot parameter optimisation in Part~2, and the algorithm is straightforward to implement from the pseudocode provided.
The publicly available MATLAB source code facilitates cross-verification.

\bibliographystyle{IEEEtran}
\begin{thebibliography}{99}

\bibitem{mirjalili2016dragonfly}
S.~Mirjalili, ``Dragonfly algorithm: a new meta-heuristic optimization technique for solving single-objective, discrete, and multi-objective problems,''
\textit{Neural Comput.\ Appl.}, vol.~27, no.~4, pp.~1053--1073, 2016.

\bibitem{kennedy1995pso}
J.~Kennedy and R.~Eberhart, ``Particle swarm optimization,'' in
\textit{Proc.\ IEEE Int.\ Conf.\ Neural Networks}, 1995, pp.~1942--1948.

\bibitem{wolpert1997nfl}
D.~H.~Wolpert and W.~G.~Macready, ``No free lunch theorems for optimization,''
\textit{IEEE Trans.\ Evol.\ Comput.}, vol.~1, no.~1, pp.~67--82, 1997.

\bibitem{reynolds1987flocks}
C.~W.~Reynolds, ``Flocks, herds and schools: a distributed behavioral model,''
\textit{ACM SIGGRAPH Comput.\ Graph.}, vol.~21, no.~4, pp.~25--34, 1987.

\bibitem{saremi2014transfer}
S.~Saremi, S.~Mirjalili, and A.~Lewis, ``How important is a transfer function in discrete heuristic algorithms,''
\textit{Neural Comput.\ Appl.}, pp.~1--16, 2014.

\bibitem{coello2004mopso}
C.~A.~C.~Coello, G.~T.~Pulido, and M.~S.~Lechuga, ``Handling multiple objectives with particle swarm optimization,''
\textit{IEEE Trans.\ Evol.\ Comput.}, vol.~8, no.~3, pp.~256--279, 2004.

\bibitem{cui2009boid}
Z.~Cui and Z.~Shi, ``Boid particle swarm optimisation,''
\textit{Int.\ J.\ Innov.\ Comput.\ Appl.}, vol.~2, pp.~77--85, 2009.

\bibitem{kadrovach2002particle}
B.~A.~Kadrovach and G.~B.~Lamont, ``A particle swarm model for swarm-based networked sensor systems,'' in
\textit{Proc.\ ACM Symp.\ Appl.\ Comput.}, 2002, pp.~918--924.

\bibitem{cui2009alignment}
Z.~Cui, ``Alignment particle swarm optimization,'' in
\textit{Proc.\ 8th IEEE Int.\ Conf.\ Cognitive Informatics}, 2009, pp.~497--501.

\end{thebibliography}

\end{document}
